Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-05-10 WholePlotTrajectories.Rmd

The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
They provide wood and multiple non-timber forest products, shelter for a diversified fauna, regulate the local climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being. 
The simultaneous increase of forests products demand and substantial climatic changes heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatened the maintenance of communities structure, composition and functioning and their underlying dynamics [@Anderson-Teixeira2013;@Sist2015].

In tropical forests communities are shaped by a constant range of disturbance that changes the abiotic environment, as the light, the heat and water fluxes, and the biotic interaction and competitive pressure.
The cornerstone of tropical forests ecology is then to understand their response to disturbance and its mechanisms [@White2001;@Chazdon2003a].
Forests response has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density.
These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Denslow2000;@Blanc2009;@Rutishauser2016].
However the response of tree species diversity, which is determinant of ecosystems productivity, stability and functioning [@Tilman2014] and would be most probably impacted by post-disturbance environmental changes [@CazzollaGatti2014], remains unclear.

In the short-term diversity dynamics demonstrated negligible or even positive impacts of disturbance on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a]. 
However, validations of the IDH in the long term remain scarce and mainly rely on the analysis of species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016].
More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and evenness, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006].
Furthermore, the functional approach accounting for species biological attributes would be insightful as it reveals species fitness and ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013].
In that respect major functional traits related to species ecology and performance were largely adopted through a relevant framework [@Diaz2005;@Villeger2008a].
The functional trait-based approach, for example, highlighted in tropical rainforests the environmental filters fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008].
This translated by shifts from "conservative" slow-growing species dealing with scarce resources, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011].
It was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].
A proper monitoring of communities response should therefore encompass taxonomic and functional diversity and composition measures to test the validity of the IDH in the long term for tropical forests, and clarify the resilience of communities evenness, composition, and also functioning.
The trajectories followed by all these facets would highlight the role of deterministic processes, like competitive exclusion or abiotic selection, and the communities' convergence maintaining intrinsinc differences in diversity and composition,which is as much insgights for future adaptive conservation strategies [@Adler2007].

Here we investigated over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed.
We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity.
Specifically, we (i) tested the validity of the IDH in the long term for tropical hiperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning (iii) questioned the completeness of communities recovery given the altered functional redundancy.
>>revoir la fin

>>>>

The large areas covered with tropical forests worldwide hold crucial economic, social and cultural value. 
They provide wood and multiple non-timber forest products, shelter for a diversified fauna, regulate the local and global climate, support the carbon, water and nutrient cycles, and ensure cultural and human well-being. 
The simultaneous increase of forests products demand and substantial climatic changes has heightened the pressure on the remaining forests [@Gibson2011a; @Morales-Hidalgo2015] and threatened the maintenance of communities structure, composition and functioning and their underlying dynamics in space and time [@Anderson-Teixeira2013;@Sist2015].

In tropical forest, ecological communities are contantly re-shaped by the natural disturbance events that changes the abiotic environment, as the light, the heat and water fluxes, and the biotic interaction and competitive pressure *(Goulamoussene Biogeosciences)*. The cornerstone of tropical forests ecology is then to understand the underlyings of ecosystem response to disturbance, its mechanisms and drivers [@White2001;@Chazdon2003a]. For now, this has been largely studied through structural parameters rapid and convenient to measure, as aboveground biomass, tree height or stem density *(e.g. PiponiotElife, RutishauserBiotropica)*. These structural parameters have then been sucessfully modeled and allowed to assess the maintenance of ecosystems processes and services [@Herault2018]. However the response of tree species diversity, which is determinant of ecosystems productivity, stability and functioning [@Tilman2014, Liang2016Science] and would be most probably impacted by post-disturbance environmental changes [@BaralotoJAE], remains unclear.

In the short-term diversity dynamics demonstrated negligible or even positive impacts of disturbance on communities diversity, which have been formalized by the intermediate disturbance hypothesis (IDH) stating a maximized species diversity at intermediate disturbance intensity [@Molino2001;@Kariuki2006a;@Berry2008a]. However, validations of the IDH in the long term remain scarce and mainly rely on the analysis of the rough species richness that gives limited or misleading information on forests recovery and functioning [@Martin2015;@Chaudhary2016]. More releveant monitoring would encompass communities composition, that is crucial for conservation issues, and diversity profiles, that reveals ecological rules underlying communities' structure [@Magurran1988;@Lavorel2002;@Bellwood2006]. Furthermore, the functional approach accounting for differenes in species biological attributes and functioning would be insightful as it reveals the species role in the post-disturbance trajectory of ecosystems functioning [@Violle2007b;@Moretti2009;@Baraloto2012a;@Scheiter2013]. In that respect major functional traits related to species ecology and performance were largely adopted through an integrative framework [@Diaz2016Nature, @Kunstler2016Nature]. In disturbed tropical rainforests, this framework translates into the fostering disturbance resistant species with rapid growth and efficient resources acquisition [@Molino2001;@Haddad2008] with a shift from "conservative" slow-growing species dealing with scarce resources that dominate before disturbance, to "acquisitive" fast-growing species with rapid and efficient use of abundant resources [@TerSteege2001;@Reich2014;@Herault2011] and was mirrored by shifts in key functional traits related to resource acquisition (leaf area, density and chlorophyll content, and stem specific gravity and bark thickness), tree growth and reproduction life history traits (seed mass and maximum height) [@Wright2004;@Westoby2006a;@Chave2009b].

A proper monitoring of communities response should therefore encompass taxonomic and functional diversity and composition measures to test the validity of the IDH in the long term for tropical forests, and clarify the resilience of communities evenness, composition, and also functioning. The trajectories followed by all these facets would highlight the role of deterministic processes, like competitive exclusion or abiotic selection, and the communities' convergence maintaining intrinsinc differences in diversity and composition, which is as much insights for future adaptive management strategies [@Adler2007]. Here we investigated over 30 years the response of 75 ha of forests plots set up on a gradient of disturbance intensity, from 10 to 60% of ecosystem biomass removed. We made use of a large functional traits database browsing major leaf, stem and seed traits and species maximum height to draw the trajectories over time of communities taxonomic and functional composition and diversity. Specifically, we (i) tested the validity of the IDH in the long term for tropical hyperdiverse forest and highlighted the ecological rules shaping their response to disturbance, (ii) clarified the different facets of communities resilience in terms of communities composition, diversity and functioning *????(iii) questioned the completeness of communities recovery given the altered functional redundancy.*
